Hierarchically Supervised Latent Dirichlet Allocation

### 2011-05-26 paper/discussion.tex
a first draft (sucky) abstract and introduction
Updated Experiment section and other minor things

\begin{itemize}
\item what about the nonparametric version of this?
\end{itemize}

>>>>

We have described a mixed membership model with hierarchical supervision. We have demonstrated this model in the context of document modeling with hierarchical multi-label supervision. Such a model is appropriate in domains where there are hierarchical constraints among the labels such as is the case in an IS-A hierarchy.

...

\begin{itemize}
\item what about the nonparametric version of this?
\item discuss the broader goal, from the beginning of search engine time, to combine categorization and free text.  this, to our knowledge, is the first principled approach to doing so
\end{itemize}

### 2011-05-26 paper/main.tex
Updated Experiment section and other minor things

		\item For each label $l \in \mathcal{L}$:
		\begin{itemize}
			\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$ 
			\item Set the response variable \begin{equation*} y_{l, d} \ | \ a_{l, d} = \begin{cases}
																										 \ \ \ 1 & \text{if $a_{l,d} > 0$ and $y_{{\rm parent}(l),d} = 1$} \\
																										-1 & \text{otherwise}
																								\end{cases}  
															 \end{equation*}															 
		\end{itemize}
	\end{itemize}
\end{enumerate}
%
This type of generative model is known as a probit regression model. Probit regression is like logistic regression except instead of modeling the logit of $P(y = 1)$ using a linear form we model $\Phi^{-1}(P(y = 1))$ using a linear form, where $\Phi(\cdot)$ is the CDF for a standard normal distribution.  In this case, the regression is conditional on the parents according to the constraints of the labeling hierarchy. The latent variables $a_{l,d}$ utilized here are also known as an auxiliary variables because the are introduced to make exact Gibbs sampling possible and are not of primary interest.  
%
\section{Inference}
\label{sec:inference}
%
In the Bayesian approach to statistical modeling the primary task of inference is to find the posterior distribution over the unobserved parameters of the model. However, it will often be the case that the set of labels $\mathcal{L}$ is not fully observed for every document.  We will define $\mathcal{L}_d$ to be the subset of labels which have been observed for document $d$.  It is straightforward to integrate out the variables $a_{l^\prime,d}$ and $y_{l^\prime,d}$ for $l^\prime \in \mathcal{L} \backslash \mathcal{L}_d$ from the full generative model.  We can also integrate out the parameters $\mathbf{\phi_{1:K}}$ and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}.  Therefore, in our model the latent variables are $\mathbf{z} = \{ z_{1:N_d,d} \}_{d = 1, \ldots, D},\mathbf{\eta} = \{ \mathbf{\eta_l} \}_{l \in \mathcal{L}},\mathbf{a} = \{ a_{l,d}\}_{l \in \mathcal{L}, d = 1, \ldots, D}, \mathbf{\beta}, \alpha, \alpha^{\prime}$ and $\gamma$.

@@

Posterior inference is performed using the {}``direct assignment'' method of \citet{TehJorBea2006}.
%
\begin{equation}
\beta\sim {\rm Dir}\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)
\end{equation}
%
\begin{equation}
p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}
\end{equation}
%
%
where $s\left(n,m\right)$ represents stirling numbers of the first kind.
%
\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha' \right)$, $p\left(\gamma \right)$}
%
The hyperparameters $\alpha$, $\alpha^\prime$, and $\gamma$ are given broad ${\rm Gamma}(2, 1000)$ prior distributions and sampled via the Metropolis-Hastings algorithm.
%
\subsection{Prediction}

@@

Given the expectation o

@@

\subsection{Related Work}
\label{sec:related_work}
Latent dirichlet allocation (LDA) is a generative probabilistic model
of corpora that represents documents as a mixed membership bag-of-words.
Also known as topic models, these models infer the latent structure,
or topics, of documents in a corpus. Each document is represented
as a collection of words, generated from a set of topic assignments
(one for each word), where each topic assignment is drawn from a distribution
over topics \citep{Blei2003}.

This model, supervised latent dirichlet allocation (sLDA), builds
on LDA by incorporating an exponential family response variable. Although
there are many models for making predictions based on free text, sLDA
is unique in that it is a generative model, it represents documents
as a mixed-membership, and constrains the inference of the latent
structure of the documents by its predictability of the response variable.
In other words, sLDA infers topics such that the model is capable
of a high predictive likelihood for words in a document and the response
variable associated with a document. This approach has been shown
to outperform both LASSO (L1 regularized least squares regression)
and LDA followed by least squares regression \citep{BleiMcAuliffe2008}.

Other models - predicting document links, other supervised latent
variable models

% move to related work?
While there have been some attempts to automatically classify groups
of patients as a potential preliminary step to ICD-9 code assignment
\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
automatic assignment of ICD-9 codes to medical text became a more
prevalent research topic only in the last few years. A subset of earlier
work proposed various methods on small corpora, based on a few specific
diseases \citep{Rao2003} but the most recent and promising work on
the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
(website). Most of the classification strategies included word matching
and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
The data set given to the participants consisted only of documents
that were 1-2 lines each and all of the documents were radiology reports
- clearly limiting the scope of potential ICD-9 codes which could
be assigned. The only paper which has attempted to work with a document
scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
Lita proposed support vector machine and bayesian ridge regression
methods to assign appropriate labels to the documents but did not
utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

@@

\newpage{}

\subsection{Rao-Blackwellization}

To derive the Gibbs sampler in general, we integrate over the parameters
$\theta$ and $\phi_{1:K}$ resulting in the collapsed joint distribution
shown in Equations 5-8.

\begin{landscape} \centering % optional, probably makes it look better to have it centered on the page

\begin{small}

\begin{equation}
p\left(\theta,z_{1:N}\mid w_{1:N},y_{1:I},\phi_{1:K},\eta_{1:I},\alpha,\beta,\mu,\xi\right)=\frac{p\left(\theta\mid\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)}{\int_{\theta}p\left(\theta\mid\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)d\theta}\end{equation}

\begin{equation}
\intop_{\theta}\intop_{\phi_{1:K}}p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\theta},\mathbf{\phi},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)d\theta d\phi_{1:K}=\intop_{\theta}\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\left\{ p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}\left[p\left(z_{m,n}\mid\theta_{m}\right)p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)\right]\prod_{i=1}^{I}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right\} \prod_{i=1}^{I}p\left(\eta_{i};\mu\right)d\theta d\phi_{1:K}\end{equation}

\begin{equation}
p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\prod_{n=1}^{N}p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)d\phi_{1:K}\intop_{\theta}\prod_{m=1}^{M}p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}p\left(z_{m,n}\mid\theta_{m}\right)d\theta\end{equation}

\begin{equation}
=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}\end{equation}

\begin{equation}
=\prod_{i=1}^{I}\left[\mathcal{N}\left(\eta_{i}\mid\mu,1\right)\prod_{m=1}^{M}h\left(y\right)\exp\left\{ \left(\eta_{i}^{T}\bar{z}\right)y-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}

\end{small} \end{landscape}
\end{document}

%\begin{align}
%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right) & =\\
% & \frac{p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)p\left(\theta\mid\beta,\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}{\int_{\alpha,\alpha',\gamma}p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)\int_{\theta}p\left(\theta\mid\beta,\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}\end{align}

>>>>

		\item For each label $l \in \mathcal{L}$:
		\begin{itemize}
			\item Draw $a_{l,d}\mid z_{1:N_{d},d},\eta_{l},y_{parent(l),d}\sim\begin{cases}
				\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right), & y_{parent(l)}=1\\
				\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ 
			%\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$ 
			\item Set the response variable \begin{equation*} y_{l, d} \ | \ a_{l, d} = \begin{cases}
																										 \ \ \ 1 & \text{if $a_{l,d} > 0$ and $y_{{\rm parent}(l),d} = 1$} \\
																										-1 & \text{otherwise}
																								\end{cases}  
															 \end{equation*}															 
		\end{itemize}
	\end{itemize}
\end{enumerate}
%
This type of generative model is known as a probit regression model. Probit regression is like logistic regression except instead of modeling the logit of $P(y = 1)$ using a linear form we model $\Phi^{-1}(P(y = 1))$ using a linear form, where $\Phi(\cdot)$ is the CDF for a standard normal distribution.  In this case, the regression is conditional on the parents according to the constraints of the labeling hierarchy. The latent variables $a_{l,d}$ utilized here are also known as an auxiliary variables because the are introduced to make exact Gibbs sampling possible and are not of primary interest.  
%
\section{Inference}
\label{sec:inference}
%
In the Bayesian approach to statistical modeling the primary task of inference is to find the posterior distribution over the unobserved parameters of the model. However, it will often be the case that the set of labels $\mathcal{L}$ is not fully observed for every document.  We will define $\mathcal{L}_d$ to be the subset of labels which have been observed for document $d$.  It is straightforward to integrate out the variables $a_{l^\prime,d}$ and $y_{l^\prime,d}$ for $l^\prime \in \mathcal{L} \backslash \mathcal{L}_d$ from the full generative model.  We can also integrate out the parameters $\mathbf{\phi_{1:K}}$ and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}.  Therefore, in our model the latent variables are $\mathbf{z} = \{ z_{1:N_d,d} \}_{d = 1, \ldots, D},\mathbf{\eta} = \{ \mathbf{\eta_l} \}_{l \in \mathcal{L}},\mathbf{a} = \{ a_{l,d}\}_{l \in \mathcal{L}, d = 1, \ldots, D}, \mathbf{\beta}, \alpha, \alpha^{\prime}$ and $\gamma$.

@@

Posterior inference is performed using the {}``direct assignment'' method of \citet{TehJorBea2006}.
%
\begin{equation}
\beta\sim {\rm Dir}\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)
\end{equation}
%
\begin{equation}
p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}
\end{equation}
%
%
where $s\left(n,m\right)$ represents stirling numbers of the first kind.
%
\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha' \right)$, $p\left(\gamma \right)$}
%
The hyperparameters $\alpha$, $\alpha^\prime$, and $\gamma$ are given broad ${\rm Gamma}(2, 1000)$ prior distributions and sampled via the Metropolis-Hastings algorithm.
%

@@

The two measures for predictive performance used here include the
true positive rate and the false positive rate. We evaluate model
performance on held out data. A more ideal evaluation of performance
would include a manually labeled hierarchy since it is well known
that ICD-9 codes have a relatively low sensitivity.

Performance was evaluated against $p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$.

@@

TBD

\subsection{Related Work}

\label{sec:related_work} Latent dirichlet allocation (LDA) is a generative
probabilistic model of corpora that represents documents as a mixed
membership bag-of-words. Also known as topic models, these models
infer the latent structure, or topics, of documents in a corpus. Each
document is represented as a collection of words, generated from a
set of topic assignments (one for each word), where each topic assignment
is drawn from a distribution over topics \citep{Blei2003}.

Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
an exponential family response variable. Although there are many models
for making predictions based on free text, sLDA is unique in that
it is a generative model, it represents documents as a mixed-membership,
and constrains the inference of the latent structure of the documents
by its predictability of the response variable. In other words, sLDA
infers topics such that the model is capable of a high predictive
likelihood for words in a document and the response variable associated
with a document. This approach has been shown to outperform both LASSO
(L1 regularized least squares regression) and LDA followed by least
squares regression \citep{BleiMcAuliffe2008}.

There have been many models that incorporate both latent models of
text and some form of supervision {[}citations{]}. One set of models
that are particularly relevant to HSLDA are Chang and Blei's {[}citation{]}
hierarchical models for document networks. In that family of models,
they encountered a similar scenario where the lack of a link did not
truly indicate absence. In our model, the lack of a code in the hierarchy
being assigned does not necessarily indicate absence. Therefore, as
in the work of Chang and Blei, we employ regularization in the form
of a negative prior on the regression parameters to provide a prior
that indicates a bias towards being truly negative in the absence
of a code.

% move to related work?
While there have been some attempts to automatically classify groups
of patients as a potential preliminary step to ICD-9 code assignment
\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
automatic assignment of ICD-9 codes to medical text became a more
prevalent research topic only in the last few years. A subset of earlier
work proposed various methods on small corpora, based on a few specific
diseases \citep{Rao2003} but the most recent and promising work on
the subject was inspired by the 2007 Medical NLP Challenge: {}``International
Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
(website). Most of the classification strategies included word matching
and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
The data set given to the participants consisted only of documents
that were 1-2 lines each and all of the documents were radiology reports
- clearly limiting the scope of potential ICD-9 codes which could
be assigned. The only paper which has attempted to work with a document
scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
Lita proposed support vector machine and bayesian ridge regression
methods to assign appropriate labels to the documents but did not
utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

@@

\end{document}
